---
title: "Assign FAO Areas to stocks in the RAM Legacy Stock Assessment Database (v4.491)"
output: html_document
---

```{r setup, include=FALSE}
# Load packages
library(knitr) 
library(countrycode) # country name matching
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)

# Path to raw RAM .Rdata
ram_dat_path <- "./RAM v4.491 Files (1-14-20)/RAM v4.491/DB Files With Assessment Data/R Data/DBdata[asmt][v4.491].RData"

# Load RAM data
load(ram_dat_path)

# Get relevant metadata to start assigning FAO regions
ram_stocks_spatial_areas <- stock %>%
  dplyr::select(stockid, 
                stocklong,
                areaid, 
                region) %>%
  left_join(area %>% dplyr::select(areaid, areaname, country), by = "areaid") %>%
  left_join(metadata %>% dplyr::select(stockid, primary_FAOarea), by = "stockid") %>%
  arrange(country)

```

## Assign FAO areas based on "region"

Though the RAM database now has a "primary FAO area" attribute associated with many of the entries, it's worth making our own lookup table to make sure they match up. We will also want to correct these for the multinational stocks.

We start by using the "region" attribute. For some, this will be enough to determine the relevant FAO area. For others it will not, and we will need to assign the relevant FAO area based on another arribute. 

```{r cars}
ram_fao_by_region <- ram_stocks_spatial_areas %>%
  distinct(region) %>%
  mutate(region_fao_area = case_when(region == "South America" ~ "NA", # Not sufficienct to determine FAO area
                                     region == "Australia" ~ "NA", # Not sufficienct to determine FAO area
                                     region == "Canada East Coast" ~ "21",
                                     region == "Canada West Coast" ~ "67",
                                     region == "Canada West Coast (Pacific Salmon)" ~ "2",
                                     region == "Other" ~ "NA", # Not sufficienct to determine FAO area
                                     region == "Japan" ~ "61",
                                     region == "Russia Japan (Pacific Salmon)" ~ "4",
                                     region == "Indian Ocean" ~ "NA", # Not sufficienct to determine FAO area
                                     region == "Mediterranean-Black Sea" ~ "37",
                                     region == "Atlantic Ocean" ~ "NA", # Not sufficienct to determine FAO area
                                     region == "Pacific Ocean" ~ "NA", # Not sufficienct to determine FAO area
                                     region == "European Union" ~ "NA", # Not sufficienct to determine FAO area
                                     region == "West Africa" ~ "NA", # Not sufficienct to determine FAO area
                                     region == "Europe non EU" ~ "NA", # Not sufficienct to determine FAO area
                                     region == "US Alaska" ~ "NA", # Not sufficienct to determine FAO area
                                     region == "New Zealand" ~ "81",
                                     region == "South Africa" ~ "NA", # Not sufficienct to determine FAO area
                                     region == "US East Coast" ~ "NA", # Not sufficienct to determine FAO area
                                     region == "US West Coast" ~ "NA", # Not sufficienct to determine FAO area
                                     region == "US Southeast and Gulf" ~ "31",
                                     region == "US Alaska (Pacific Salmon)" ~ "2",
                                     region == "US West Coast (Pacific Salmon)" ~ "2",
                                     TRUE ~ "NA")) %>%
  dplyr::filter(region_fao_area != "NA")
```

## Assign FAO areas based on "country"

For instances where region was not enough, we now consider country. 

```{r}
ram_fao_by_country <- ram_stocks_spatial_areas %>%
  anti_join(ram_fao_by_region, by = "region") %>%
  distinct(country) %>%
  mutate(country_fao_area = case_when(country == "Argentina" ~ "41",
                                      country == "Australia" ~ "NA", # Not sufficienct to determine FAO area
                                      country == "Chile" ~ "87",
                                      country == "Iran" ~ "4",
                                      country == "multinational" ~ "NA", # Not sufficienct to determine FAO area
                                      country == "Multinational" ~ "NA", # Not sufficienct to determine FAO area
                                      country == "Peru" ~ "87",
                                      country == "Russia" ~ "NA", # Not sufficienct to determine FAO area
                                      country == "South Africa" ~ "NA", # Not sufficienct to determine FAO area
                                      country == "USA" ~ "NA", # Not sufficienct to determine FAO area
                                      TRUE ~ "NA")) %>%
  dplyr::filter(country_fao_area != "NA")
```

## Assign FAO areas based on "areaname"

For instances where region and country were not enough, we now consider "areaname". Let's break it apart by country to make it a bit more feasible. In some cases it's necessary to reference the specific stock names to better assign regions. 

### Australia

```{r}
ram_fao_by_areaname_australia <- ram_stocks_spatial_areas %>%
  anti_join(ram_fao_by_region, by = "region") %>%
  dplyr::filter(country == "Australia") %>%
  distinct(country, region, areaname, stocklong) %>%
  mutate(areaname_fao_area = case_when(areaname == "New South Wales to Western Australia" ~ "57;81",
                                       areaname == "Southeast Australia" ~ "57;81",
                                       areaname == "Northern Australia" ~ "57;71",
                                       areaname == "Eastern half of Southeast Australia" ~ "81",
                                       areaname == "Western half of Southeast Australia" ~ "57",
                                       areaname == "Queensland and New South Wales" ~ "71;81",
                                       areaname == "Cascade Plateau" ~ "81", # might include a small part of 57
                                       areaname == "Macquarie Island" ~ "81",
                                       areaname == "Eastern Australia" ~ "57",
                                       areaname == "South Australia Northern Zone" ~ "57",
                                       areaname == "South Australia Southern Zone" ~ "57",
                                       areaname == "Northern Spencer Gulf" ~ "57",
                                       areaname == "Southern Gulf St. Vincent" ~ "57",
                                       areaname == "Southern Spencer Gulf" ~ "57",
                                       areaname == "Tasmania" ~ "57",
                                       TRUE ~ "57;71;81")) # For future/historical stocks, do all three
```

### Multinational

This one will need further splitting apart

```{r}
multinational_countries <- ram_stocks_spatial_areas %>%
  anti_join(ram_fao_by_region, by = "region") %>%
  dplyr::filter(country %in% c("multinational", "Multinational")) %>%
  distinct(region, country)
```

#### Europe

```{r}
ram_fao_by_areaname_multinational_europe <- ram_stocks_spatial_areas %>%
  anti_join(ram_fao_by_region, by = "region") %>%
  dplyr::filter(country %in% c("multinational", "Multinational")) %>%
  dplyr::filter(region %in% c("Europe non EU", "European Union")) %>%
  distinct(country, region, areaname, stocklong) %>%
  mutate(areaname_fao_area = case_when(areaname == "Azores" ~ "27;34", # crosses both
                                       areaname == "NAFO 1f and ICES 14" ~ "21;27", # crosses both
                                       areaname == "ICES 5-12-14 and NAFO Subareas 1-2 (deep)" ~ "21;27",
                                       areaname == "ICES 5-12-14 and NAFO Subareas 1-2 (shallow)" ~ "21;27",
                                       areaname == "Azores Grounds" ~ "27;34",
                                       areaname == "Mid-Atlantic Ridge" ~ "27", # check on this one
                                       TRUE ~ "27")) # Otherwise we assume it's 27
```

#### South America

```{r}
ram_fao_by_areaname_multinational_south_america <- ram_stocks_spatial_areas %>%
  anti_join(ram_fao_by_region, by = "region") %>%
  dplyr::filter(country %in% c("multinational", "Multinational")) %>%
  dplyr::filter(region %in% c("South America")) %>%
  distinct(country, region, areaname, stocklong) %>%
  mutate(areaname_fao_area = case_when(areaname == "Chile" ~ "87",
                                       areaname == "Chilean EEZ and offshore" ~ "87",
                                       areaname == "South Chile Argentina" ~ "41;87",
                                       TRUE ~ "41;87"))

```

#### West Africa
```{r}
ram_fao_areaname_multinational_west_africa <- ram_stocks_spatial_areas %>%
  anti_join(ram_fao_by_region, by = "region") %>%
  dplyr::filter(country %in% c("multinational", "Multinational")) %>%
  dplyr::filter(region %in% c("West Africa")) %>%
  distinct(country, region, areaname, stocklong) %>%
  mutate(areaname_fao_area = case_when(areaname == "West Africa" ~ "34", # small pelagics and crustaceans
                                       areaname == "West Africa Subgroup 1" ~ "34",
                                       areaname == "West Africa Subgroup 2" ~ "34",
                                       areaname == "West Africa Subgroup 3" ~ "34;47",
                                       areaname == "West Africa Zone A+B" ~ "34",
                                       areaname == "West Africa Zone C" ~ "34",
                                       TRUE ~ "34"))
```

#### US Alaska

```{r}
ram_fao_areaname_multinational_alaska <- ram_stocks_spatial_areas %>%
  anti_join(ram_fao_by_region, by = "region") %>%
  dplyr::filter(country %in% c("multinational", "Multinational")) %>%
  dplyr::filter(region %in% c("US Alaska")) %>%
  distinct(country, region, areaname, stocklong) %>%
  mutate(areaname_fao_area = case_when(areaname == "North Pacific" ~ "61;67", 
                                       TRUE ~ "61;67"))
```

#### Ocean basins and other

```{r}
ram_fao_by_areaname_multinational_oceans <- ram_stocks_spatial_areas %>%
  anti_join(ram_fao_by_region, by = "region") %>%
  dplyr::filter(country %in% c("multinational", "Multinational")) %>%
  dplyr::filter(region %in% c("Atlantic Ocean", "Indian Ocean", "Pacific Ocean", "Other")) %>%
  distinct(country, region, areaname, stocklong) %>%
  mutate(areaname_fao_area = case_when(areaname == "Indian Ocean" ~ "51;57",
                                       areaname == "Northern Atlantic" ~ "21;27;31;34", 
                                       areaname == "South Atlantic" ~ "34;41;47", 
                                       areaname == "South Pacific Ocean	" ~ "71;77;81;87", 
                                       areaname == "Eastern Atlantic" & stocklong == "Atlantic bluefin tuna Eastern Atlantic" ~ "27;37;34;47",
                                       areaname == "Eastern Atlantic" ~ "27;34;47",
                                       areaname == "Western Atlantic" ~ "21;31;41",
                                       areaname == "Ross Sea" ~ "88",
                                       areaname == "Atlantic Ocean" ~ "21;27;31;34;41;47", 
                                       areaname == "Central Western Pacific Ocean" ~ "61;67;71;77;81", 
                                       areaname == "Eastern Pacific" ~ "67;77;81;87",
                                       areaname == "Southern Oceans" ~ "48;58;88", # Actually extends up higher
                                       areaname == "Nothwest Pacific Ocean" ~ "61",
                                       areaname == "Northeast Pacific" ~ "77;87",
                                       areaname == "Western Pacific Ocean" ~ "71;81", # should be SW
                                       areaname == "North Pacific Ocean" ~ "61;67;71;77",
                                       areaname == "Pacific Ocean" ~ "61;67;71;77;81;87",
                                       areaname == "Western and Central North Pacific" ~ "61;71;67;77",
                                       areaname == "South Pacific Ocean" ~ "71;77;81;87",
                                       
                                       TRUE ~ "NA"))
```

#### Combine Multinational 

```{r}
ram_fao_by_areaname_multinational <- ram_fao_by_areaname_multinational_europe %>%
  bind_rows(ram_fao_by_areaname_multinational_south_america) %>%
  bind_rows(ram_fao_areaname_multinational_west_africa) %>%
  bind_rows(ram_fao_areaname_multinational_alaska) %>%
  bind_rows(ram_fao_by_areaname_multinational_oceans)
```

### Russia

```{r}
ram_fao_by_areaname_russia <- ram_stocks_spatial_areas %>%
  anti_join(ram_fao_by_region, by = "region") %>%
  dplyr::filter(country == "Russia") %>%
  distinct(country, region, areaname, stocklong) %>%
  mutate(areaname_fao_area = "61")
```

### South Africa

```{r}
ram_fao_by_areaname_south_africa <- ram_stocks_spatial_areas %>%
  anti_join(ram_fao_by_region, by = "region") %>%
  dplyr::filter(country == "South Africa") %>%
  distinct(country, region, areaname, stocklong) %>%
  mutate(areaname_fao_area = case_when(areaname == "South Africa Areas 1-2" ~ "47",
                                       areaname == "South Africa Areas 3-4" ~ "47",
                                       areaname == "South Africa Areas 5-6" ~ "47",
                                       areaname == "South Africa Area 7" ~ "47",
                                       areaname == "South Africa Area 8" ~ "47",
                                       areaname == "South Africa South coast" ~ "47",
                                       areaname == "South Africa Subantarctic Prince Edward Islands" ~ "51;58",
                                       areaname == "South Africa" ~ "47;51",
                                       TRUE ~ "47;51"))
```

### USA

```{r}
ram_fao_by_areaname_usa <- ram_stocks_spatial_areas %>%
  anti_join(ram_fao_by_region, by = "region") %>%
  dplyr::filter(country == "USA") %>%
  distinct(country, region, areaname, stocklong) %>%
  mutate(areaname_fao_area = case_when(areaname == "Gulf of Maine / Georges Bank" ~ "21",
                                       areaname == "Bering Sea and Aleutian Islands" ~ "61;67",
                                       areaname == "Gulf of Alaska" ~ "67",
                                       areaname == "Pacific Coast" ~ "67;77", # check
                                       areaname == "Mid-Atlantic Coast" ~ "21;31", # check
                                       areaname == "Chesapeake Bay" ~ "21",
                                       areaname == "Pribilof Islands" ~ "67",
                                       areaname == "Saint Matthews Island" ~ "67",
                                       areaname == "California" ~ "67;77", # barely goes into 67
                                       areaname == "Northern Pacific Coast" ~ "67", # check
                                       areaname == "Oregon Coast" ~ "67",
                                       areaname == "Southern Pacific Coast" ~ "77",
                                       areaname == "Washington" ~ "67",
                                       areaname == "Atlantic Coast" ~ "21;31",
                                       areaname == "Gulf of Maine / Georges Bank-Southern New England" ~ "21",
                                       areaname == "Gulf of Maine / Cape Hatteras" ~ "21",
                                       areaname == "Northern California" ~ "67;77",
                                       areaname == "Southern California" ~ "77",
                                       areaname == "Central Pacific Coast" ~ "77",
                                       areaname == "Georges Bank" ~ "21",
                                       areaname == "Gulf of Maine" ~ "21",
                                       areaname == "Northwestern Atlantic Coast" ~ "21",
                                       areaname == "Prince William Sound" ~ "67",
                                       areaname == "Sitka" ~ "67",
                                       areaname == "Togiak District" ~ "67",
                                       areaname == "Southern New England" ~ "21",
                                       areaname == "Atlantic" ~ "21;31",
                                       areaname == "Gulf of Maine / Northern Georges Bank" ~ "21",
                                       areaname == "Southern Georges Bank / Mid-Atlantic" ~ "21", # check
                                       areaname == "Eastern Bering Sea and Aleutian Islands" ~ "61;67",
                                       areaname == "Aleutian Islands" ~ "61;67",
                                       areaname == "Bering Sea" ~ "61;67",
                                       areaname == "Northwestern Atlantic" ~ "21", # check
                                       areaname == "Bristol Bay" ~ "67",
                                       areaname == "Norton Sound" ~ "67",
                                       areaname == "Eastern Bering Sea / Aleutian Islands / Gulf of Alaska" ~ "61;67",
                                       areaname == "Georges Bank and Mid-Atlantic Bight" ~ "21",
                                       areaname == "North Pacific" ~ "67",
                                       areaname == "Southern New England /Mid Atlantic" ~ "21", #check
                                       areaname == "Bogoslof" ~ "67",
                                       areaname == "Eastern Bering Sea" ~ "61;67",
                                       areaname == "Alaska" ~ "61;67",
                                       areaname == "Cape Cod / Gulf of Maine" ~ "21",
                                       TRUE ~ "NA"))
```

### Combine countries

```{r}
ram_fao_by_areaname <- ram_fao_by_areaname_australia %>%
  bind_rows(ram_fao_by_areaname_multinational) %>%
  bind_rows(ram_fao_by_areaname_russia) %>%
  bind_rows(ram_fao_by_areaname_south_africa) %>%
  bind_rows(ram_fao_by_areaname_usa)
```

## Apply to stock list 

```{r}
ram_stocks_spatial_areas_out <- ram_stocks_spatial_areas %>%
  left_join(ram_fao_by_region, by = "region") %>%
  left_join(ram_fao_by_country, by = "country") %>%
  left_join(ram_fao_by_areaname, by = c("region", "country", "areaname", "stocklong")) %>%
  mutate(fao_area = case_when(!is.na(region_fao_area) ~ region_fao_area,
                              !is.na(country_fao_area) ~ country_fao_area,
                              TRUE ~ areaname_fao_area)) %>%
  dplyr::select(stockid, fao_area)

write_csv(ram_stocks_spatial_areas_out, "fao_areas_for_ram_stocks_v4.491.csv")
```

